%! Author = breandan
%! Date = 6/7/21

% Preamble
\documentclass[11pt]{article}

% Packages
\usepackage{amsmath}
\usepackage{unicode-math}
\usepackage{amssymb}

\usepackage{float}
\usepackage{tikz}

% Document
\title{Autobiographical Statement}
\author{Breandan Considine}
\begin{document}
\maketitle
I was born July 20th, 1990 in Hanover, New Hampshire. As an only child, my parents divorced at an early age, and I was raised primarily by my maternal grandparents in a town called Bethlehem. My mother worked as a bank examiner and had to travel often for work. My grandmother took care of me while my mother was away, and taught me how to pray. Some of my earliest memories were praying the rosary with her before bed. She was a catechism teacher for the local parish and together with my grandfather and uncle, made the home where I grew up.

At the age of eight, I was transferred from the public school in Bethlehem to a Catholic school in St. Johnsbury, Vermont. The school was a half an hour from home, and my grandmother would drive an hour twice a day to drop me off and pick me up after school. Later, a commuter bus started between St. Johnsbury and Littleton, and I would ride the bus with the workers to school. I was an avid reader, and loved to draw.

After graduating eighth grade, I was enrolled in a private school, St. Johnsbury Academy, where I excelled in languages. I studied Latin and Chinese and took an interest in Java programming and robotics. One of my earliest programming experiences was using a TI-83 calculator to generate prime numbers using the Sieve of Eratosthenses. I fell in love with programming, which became a lifelong passion.

After high school, I went on to the Rochester Institute of Technology, where I studied computer science and participated in the sailing club. During the summers, I worked at a small company in New Hampshire doing data analysis. I also spent year studying abroad in Shanghai, China. One day while attending Mass there, I ran into an old classmate who, unbeknownst to me, was studying there at the same time.

This brief experience has reverberated in different times and similar ways over the years, but this was my first memorable encounter with the Holy Spirit. That and similar experiences throughout my life have reminded me that He is alive and at work in the world.

After graduating from college, I worked as an engineer developing automated bidding software at an ad tech startup in Austin, Texas. Following that job, I worked for two years as a technical evangelist for a software engineering company, where I had the opportunity to travel to conferences around the world teaching programming and software engineering. I gave talks about speech recognition, developer tools and machine learning.

I returned to graduate school in 2017 to pursue machine learning, which was making significant strides in research and industry. After being admitted to the Masters degree program at University of Montreal, I helped automate the infrastructure for an AI driving competition and developed software tools for machine learning. Afterwards, I continued as a Ph.D. student at McGill University to work on machine learning for software engineering.

During my graduate studies, I had the opportunity to teach and loved working with students on research problems. I enjoyed the social aspect of organizing events together with my colleagues, through academic seminars, workshops and conferences. As a teaching assistant for several graduate courses in machine learning, I enjoyed working in fast-paced learning environments with motivated people on challenging problems.

Another one of the joys I experienced during graduate school was the joy of understanding. As a researcher, you spend years stumbling around in the dark, cursing and toiling over some wretched problem, chasing some elusive idea through dark and dingy alleys. Then, when you least expect: a flash of inspiration - lightning strikes! Research, for me, is long periods of doubt punctuated by fleeting moments of clarity. Like meeting an old friend in a strange land, or unraveling a clue which has long haunted you but failed to reveal its meaning. These vivid moments make all the confusion, doubt and toil worthwhile and worth a hundredfold more.

One of the difficulties I have faced as a researcher is keeping pride at bay. Academia celerbrates intellectual achievement -- worships it even. I have seen firsthand what remarkable things the mind is capable of when properly motivated. This encourages healthy competition and some degree of honesty, but -- I now realize -- can feed the ego and lead to false idols.

During the course of pursuing higher education, I lost my connection to the faith. I stopped attending Mass, and began to worship at the altar of academia. With the advent of intelligent machines, I began to question what humans are meant to do here on earth. After this dark period, I have come to realize the truth of what I was taught long ago: we are born sinners, and redeemed by Christ's sacrifice. We are called to follow His example, by giving of ourselves to others. Neither artificial minds, nor sin, nor forbidden fruit from the tree of knowledge changes this.

I first heard the call to religious vocation when I began to realize all these experiences are meaningful: what we are seeking is not epiphany, serendipity or insight, but the kingdom of heaven itself. Glimpses from across the shroud, of another world. Like Matthew the tax collector, who meets Jesus of Nazereth, leaves his posessions behind and follows him. Or the merchant (Matthew 13:45-46), who upon finding the pearl of great price, sells all his belongings to purchase it. Or Ambraham, who hears God's voice and offers up his only son in sacrifice. What is ours is not ours, but given by God, and to Him we must freely give in return.

As I reflect on my life, I am conscious of the tremendously fortunate circumstances I have enjoyed, enabled by the sacrifices of others. I am filled with gratitude for the opportunity to pursue my interests, and troubled by the enormous debt of goodwill I have amassed, which I deserve not. I feel a deep responsibility to spend the remainder of my life repaying this debt, and a strange but growing enthusiasm for the difficult journey that lies ahead.

For this reason, I am applying to the seminary to answer the call and to become a tool in His service. I am moved to praise the Lord God Almighty, to teach, and to give back to those in need. I have some talent for programming, writing and research. If God sees fit, I hope to use my skills as a teacher or mentor in some capacity and would be glad to apply myself to any role involving pedagogy or using technology to help others.

During a recent visit to Moreau Seminary, I had the chance to meet with Fr. DeRiso and Fr. Gallagher where I expressed to them that I had received God's call. Frankly, I am somewhat mystified that He would call me, a computer science student. But if God wills it, I will go. Thank you for considering my application.
\end{document}