%! Author = breandan
%! Date = 2/28/20
% https://www.cs.uoregon.edu/research/summerschool/summer22/registration.php

\documentclass{letter}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{amsmath}
\address{Drs. Jin Guo and Xujie Si\\McGill University\\3480 Rue University\\Montr\'eal, Qu\'ebec H3A 0E9}
\begin{document}
    \begin{letter}{Direction de l'enregistrement et de l'\'evaluation comparative\\
    Minist\`ere de l'Immigration, de la Francisation et de l'Int\'egration\\
    285, rue Notre-Dame Ouest, $4^\text{e}$ \'etage\\
    Montr\'eal, Qu\'ebec H2Y 1T8}
        \opening{Dear Immigration Officer,}
\vspace{10pt}
        We, the undersigned professors at McGill University, are writing in support of Breandan Considine, a doctoral student under our joint supervision, who is applying to renew his study permit. This letter is to confirm that Mr. Considine is currently in good standing with the university, certify his academic progress, and attest to his ability to complete the doctoral program in a timely manner.

        Since starting his Ph.D. three years ago, Breandan has successfully passed his comprehensive examination, had one peer-reviewed workshop paper accepted, and is currently working on another paper that will be submitted to a conference later this year. During that time, Breandan was also employed as a graduate teaching assistant, been nominated for and received a departmental TA award, and independently organized multiple workshops at international conferences.

        Breandan is supported by an annual stipend of \$25,000 CAD, which he will continue to receive while he remains in good standing. Considering Breandan's ongoing commitment to his research, we are confident in his ability to make significant contributions to the field of computer science and if approved, believe he will be able to complete his Ph.D. studies within the next three years. We therefor strongly recommend him for a three-year extension of his study permit.
        \vspace{10pt}

        \closing{Sincerely,\\
        \vspace{30pt}
        \fromname{Dr. Jin Guo \hspace{10pt} and \hspace{10pt} Dr. Xujie Si\phantom{vvv}}
        }
    \end{letter}
\end{document}